Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-01-26 Ben_Manuscripts/structure_paper/Draft.tex
new figures, better intro and more organization

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{setspace}
\usepackage{color}
\doublespacing

\title{Understanding the nanoscale structure of hexagonal phase
lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble
\and Joe Yelk \and Matthew Glaser \and Xunda Feng \and Michael R.
Shirts}

@@

  \bibliographystyle{ieeetr}
  \graphicspath{{./figures/}}
  \maketitle

@@

  Reverse osmosis (RO) and nanofiltration (NF) are membrane based separation
  processes widely used to create potable or reusable water by removing salts and
  other dissolved compounds from water sources. RO membranes are typically thin
  film composite (TFC) membranes with a porous support layer and a dense
  crosslinked polymer active layer which near indiscriminately separates all solutes.
  NF membranes are porous with pores on the order of 1 $\nm$ in size which separate 
  primarily based on solute size.

  RO and NF are finding applications beyond primarily water-based
  separation. Recovery of valued products from complex aqueous and organic waste
  is of increasing interest. This requires highly selective membranes.

@@

  Selective separation by a semipermeable membrane barrier is governed by two
  steps : barrier entry and transport through the barrier. A solute's ability to
  enter a membrane barrier is decided by its size, solubility and charge. A solute's
  size, shape and polarity affect its ability to travel across the barrier. Undesired
  solutes may still be able to enter the membrane making their diffusion rates
  relative to desired solutes important. It is important to design the membrane in response
  to desired and undesired solutes is thus important.

  NF -- pore polydispersity
  RO -- r0andom walk through polymer voids

  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control pore architecture at the molecular scale,
  thereby permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. One can acheive most membrane-based aqueous separations of
  small molecules using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

@@

  Commercial RO membranes are typically dense thin-film composite (TFC)
  membranes with a porous support layer and an active layer made of a polymer
  matrix formed through interfacial polymerization \cite{jeong_interfacial_2007}.
  During interfacial polymerization, a step-growth polymerization reaction occurs
  at the interface of an aqueous monomer solution that has been soaked into the
  support layer, and an organic solution of a second monomer. The polymer matrix
  is dense and highly cross-linked which gives high rejections but low
  permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically porous membranes made by a phase-inversion
  process. The most widely used phase-inversion process is immersion
  precipitation, during which one submerges a polymer, dissolved in a solvent, in
  a non-solvent. A solid, porous polymer membrane is all that remains once all
  solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
  size, which hurts membrane selectivity. A second technique used to create NF
  membranes is call track-etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
  however, the membranes have a low porosity and subsequently low permeability.

@@

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanoporous
  membranes with high selectivity, permitted by a precisely controlled pore size,
  and high permeability, allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Development of nanostructured porous materials has been limited by the
  ability to synthesize and scale various fundamentally sound technologies.
  Graphene sheets are atomically thick which results in excellent water
  permeability but defects during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of
  carbon nanotubes show promise \cite{humplik_nanostructured_2011}, but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm
  pores with a narrow pore size distribution, and MD simulations of these
  materials show that they exhibit complete rejection of solvated ions
  \cite{murad_molecular_1998}- however, experimental rejection was low and
  attributed to interstitial defects formed during membrane synthesis
  \cite{li_desalination_2004}. Consequently, there is a need for a scalable
  nanostructured porous membrane.

  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
  are suitable candidates for aqueous separation applications. LLCs are
  amphiphilic molecules that share the characteristic ability of nanostructured
  porous membrane materials to create highly ordered nanostructures with the
  added benefits of low cost and synthetic techniques feasible for large-scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
  wt\% added water results in formation of the lyotropic H\textsubscript{II}
  phase. In both cases, Na-GA3C11 monomers assembles into mesophases made of
  hexagonally packed, uniform size, cylinders with hydrophilic head groups
  oriented inward towards the cylinder center. This LLC assembly is then
  polymerized in situ to form a cross-linked polymer network to stabilize the
  structure. The hydrophilic region can act as a pore for aqueous separations
  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
  specific separations by changing the monomer chemistry
  \cite{resel_h2-phase_2000}.

  Research into polymerized LLC membranes has been revived in recent years.
  During early stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  one can align the mesophases using a magnetic field with subsequent
  cross-linking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it
  was shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in
  diameter because the pores are too large \cite{zhou_supported_2005}. We do not
  yet understand how to controllably reduce the effective pore size or how to
  tune the chemical environment in the nanopores of this or related materials for
  effective water desalination and small organic molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. It will be challenging to efficiently narrow down
  the large design space in a laboratory setting without a robust model.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group may influence pore size for size exclusion driven separations. We
  can interchange counterions which may influence both the pore size and the
  strength of the Donnan potential which affects the degree to which the membrane
  can exclude charged species.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the estimated volume of the liquid crystal
  monomers proposes that each layer consists of seven monomers
  \cite{resel_structural_2000}. Our best chance to answer this question is by
  using a molecular model orders of magnitude larger than any other reported
  atomistic liquid crystal membrane simulation, as we present here. We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  Strong $\pi$-$\pi$ stacking interactions in the direction perpendicular to
  the membrane plane support the existence of layers experimentally. $\pi$-$\pi$
  stacking will only occur between the aromatic monomer head groups which leaves
  no description of what is happening in the monomer tail region. The tails may
  entangle isotropically while maintaing stacking order among headgroups.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations over a sandwiched
  conformation \cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006} In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configuration is most likely.

@@

  %\item What constitutes a pore, and how well-defined are the pore regions? \label{point:poredefinition}
  \item What constitutes a pore, and what does the pore architecture look like? \label{point:poredefinition}

@@

  While past work describes the Col\textsubscript{h} phase as dry
  ~\cite{feng_scalable_2014}, it has been suggested by experimentalists, in
  unpublished communications, that it is likely that neat monomer leaches small
  amounts of ambient water. Experimentally, achieving a hexagonal phase with a
  completely dry system has proven difficult. If neat monomer sits in ambient
  conditions, its color turns from transparent to slightly opaque and a hexagonal
  phase forms. Although we will not explore whether water is necessary for
  self-assembly, we hypothesize that the hydrogen bonding network formed by the
  water may play a role in structuring the pores and holding together the
  hexagonal phase. We can use simulated X-ray diffraction patterns to see if
  there is any meaningful structural difference between a ``dry'' and ``wet''
  system.

@@

	\item The location of the first is at $q_z$ = 1.7 \angstrom$^{-1}$,
	corresponding to a real space separation of 3.7 \angstrom~. Previous
	work~\cite{feng_scalable_2014} attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

	\item A weak intensity line, located at exactly half the $q_z$ value of
	R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponds to a real space periodic
	spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis
	\cite{feng_scalable_2014}, meaning that if one traces the positions of the
	aromatic rings with a helical curve, then for each full turn in the helix, one
	will encounter two aromatic rings. For this reason we will refer to reflection
	as R-helix.

	\item A low intensity ring located at r = 1.4 \angstrom$^{-1}$ marks
	the third major reflection of interest. The real space separation corresponds
	to 4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains \cite{mcintosh_organization_1980}. We will call this reflection
	R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity.
	Accordingly, we name these reflection R-spots. The location of all spots is $\sim 37$
	degrees from the $q_z$ axis in their respective quadrants. In many liquid
	crystal systems one can explain the spots by the tilt angle of the alkane chains
	with respect to the membrane plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane. This plane is geometrically related the distance
	between pores. The feature, which we named R-pores, is characterized by dots
	along the equatorial axis defined when $q_z$ = 0. The spacing between dots is
	indicative of the hexagonal symmetry of the packed pores. We observe the same
	information with higher resolution using SAXS (Fig.~\ref{fig:SAXS}).

@@

% Note: we should put scripts and starting configurations (not all that were
% done but the important ones) in a public repository by the time of
% submission. 
%MRS4: I would give a ``plain text'' version of the script in supplementary. BJC: referring to parameterization script
%MRS4: somewhere would need to specify things like time step tau_t, tau_p, etc., so that people could in theory replicate

@@

%MRS4: check grammar additions in this paragraph.
  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system. To bypass the slow
  self-assembly process, we use Python scripts to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

>>>>

\documentclass[journal=jpcbfk,manusciprt=article]{achemso}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{amsmath} 
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{gensymb}

\title{Understanding the nanoscale structure of hexagonal phase lyotropic
liquid crystal membranes}
\author{Benjamin J. Coscia}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Joe Yelk}
\author{Matthew Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}

@@

  \graphicspath{{./figures/}}

  \begin{tocentry}
  \end{tocentry}

@@

  Reverse osmosis (RO) and nanofiltration (NF) are membrane-based separation
  processes widely used to create potable or reusable water by removing salts and
  other dissolved compounds from water sources. RO membranes are typically thin
  film composite (TFC) membranes with a porous support layer and a dense
  crosslinked polymer active layer which near indiscriminately separates all
  solutes. NF membranes have porous architectures with pores on the order of 1 nm
  in size which separate primarily based on solute size
  \cite{van_der_bruggen_review_2003}.

  RO and NF are finding applications beyond primarily water-based
  purifications. Recovery of valued products from complex aqueous and organic
  waste is of increasing interest. For example, flowback water (FW) produced
  during hydraulic fracturing is a complex wastewater full of potentially
  valuable dissolved organic compounds \cite{dischinger_application_2017}. There
  is increasing pressure to reuse hydraulic fracturing water rather than dispose
  of it in order to reduce social and envrionmental impacts as well as cost
  \cite{theodori_hydraulic_2014}. Rather than dispose of the wastestream
  generated in the recycling process, we can instead use highly selective
  membranes in order to successfully recover the useful compounds.

@@

  Selective separation by a semipermeable membrane barrier is governed by two
  steps: barrier entry and transport through the barrier. A solute's ability to
  enter a given membrane barrier is decided by its size, shape, charge, polarity
  and how those factors combine to interact with the membrane material. The same
  properties affect the rate at which a solute travels across the barrier. Since
  undesired solutes may still enter the membrane, their diffusion rates relative
  to desired solutes are important
  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
  selectivity, it is necessary to design membranes in a way that one can tune the
  relative diffusivities of desired and undesired solutes.

  Lyotropic liquid crystals (LLCs) are amphiphilic molecules that have the
  ability to self-assemble into porous nanostructures \cite{smith_ordered_1997}
  which can be crosslinked to create mechanically strong membrane films with
  pores on the order of 1 nm in diameter \cite{zhou_supported_2005}. Unlike
  commercial NF membranes, LLC membrane pores are uniform in size. Since LLC
  membranes lack a pore size distribution, they inherently exhibit high
  selectivity due to their strict molecular weight cut-off (MWCO).  Additionally,
  the LLCs chosen to make membranes for this application are salts.  The membrane
  gains a net surface charge when counterions from head groups that line the pore
  walls escape into the feed solution in order to balance the gradients of
  concentration and electric potential. This effect, which causes the membrane to
  repel incoming charged species, is called Donnan exclusion
  \cite{donnan_theory_1995}.

  The feasiblity of LLC membranes has been demonstrated using LLCs that form
  the type 1 bicontinuous cubic (Q\textsubscript{I})
  \cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012},
  and the inverted hexagonal (H\textsubscript{II}) \cite{zhou_supported_2005}
  phases (See Figure~\ref{fig:bcc_v_hII}). When separating organic solutes from
  NaCl, Q\textsubscript{I} phase membrane filtration experiments have shown
  selectivity 2-3 times higher than commercial RO and 6-12 times higher than
  commercial NF membranes \cite{dischinger_application_2017}.  When separating a
  series of various sized dyes, the H\textsubscript{II} phase membrane showed
  complete rejection of dyes bigger than 1.2 nm in size
  \cite{zhou_supported_2005}.

  We will focus our current efforts around the H\textsubscript{II} phase since
  its geometry should give optimal transport performance compared to the
  Q\textsubscript{I} phase. The H\textsubscript{II} phase forms at room
  temperature in the presence of c.a.~10 wt\% water and consists of hexagonally
  packed, straight, hydrophilic pore columns\cite{smith_ordered_1997}. In the
  absence of water, neat monomer will form the same hexagonal columnar structure
  which, in literature, has been referred to as the Col\textsubscript{h}
  thermotropic phase\cite{feng_scalable_2014} (See Figure~\ref{fig:hII_phase}).
  These hexagonal columnar phases represent the ideal geometry for membrane
  transport. The most promising Q\textsubscript{I} phase for membrane
  applications forms at 70\degree C when monomer is mixed with c.a.~20 wt\%
  glycerol\cite{carter_glycerol-based_2012}. The Q\textsubscript{I} phase
  consists of a network of three dimensionally interconnected pores (See
  Figure~\ref{fig:bcc_phases}). This complex geometry effectively adds a degree
  of tortuosity to the membrane which prevents optimal through-plane transport.
  Despite the geometric differences between the Q\textsubscript{I} and
  H\textsubscript{II} phase, the chemical environment within the pores is
  expected to be similar, and either membrane should be highly selective.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
        \includegraphics[height=4.5cm]{bcc_phases.png}
        \caption{}\label{fig:bcc_phases}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
        \includegraphics[height=4.5cm]{hexagonal_packing.png}
        \caption{}\label{fig:hII_phase}
  \end{subfigure}
  \caption{The choice of LLC monomer leads to different favorable liquid
        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
        represent the interface between the hydrophilic and hydrophobic regions of unit
        cell. As pictured, we assume that each region occupies an equal volume. The
        monomer shown in (b) will form the hexagonal columnar phase
        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
         water is present).}\label{fig:bcc_v_hII}
  \end{figure}

  %Nanostructured porous membrane materials have become increasingly popular for
  %aqueous separations applications such as desalination and biorefinement because
  %they offer the ability to control pore architecture at the molecular scale,
  %thereby permitting the design of solute-specific separation membranes
  %\cite{humplik_nanostructured_2011}. One can acheive most membrane-based aqueous separations of
  %small molecules using reverse osmosis (RO) or nanofiltration
  %(NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  %advances in the past few decades, they are far from perfect separation

@@

  %Commercial RO membranes are typically dense thin-film composite (TFC)
  %membranes with a porous support layer and an active layer made of a polymer
  %matrix formed through interfacial polymerization \cite{jeong_interfacial_2007}.
  %During interfacial polymerization, a step-growth polymerization reaction occurs
  %at the interface of an aqueous monomer solution that has been soaked into the
  %support layer, and an organic solution of a second monomer. The polymer matrix
  %is dense and highly cross-linked which gives high rejections but low
  %permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer

%  NF membranes are typically porous membranes made by a phase-inversion
%  process. The most widely used phase-inversion process is immersion
%  precipitation, during which one submerges a polymer, dissolved in a solvent, in
%  a non-solvent. A solid, porous polymer membrane is all that remains once all
%  solvent has been removed by non-solvent exchange
%  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
%  size, which hurts membrane selectivity. A second technique used to create NF
%  membranes is call track-etching in which a polymer film is bombarded with
%  charged particles, then chemically etched to create pores
%  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
%  however, the membranes have a low porosity and subsequently low permeability.

@@

%  The permeability-selectivity tradeoff has the potential to be overcome by
%  designing membranes at the molecular level. Next-generation nanoporous
%  membranes with high selectivity, permitted by a precisely controlled pore size,
%  and high permeability, allowed by its porous architecture, have the potential to
%  replace traditional RO and NF membrane technologies.

@@

%  Development of nanostructured porous materials has been limited by the
%  ability to synthesize and scale various fundamentally sound technologies.
%  Graphene sheets are atomically thick which results in excellent water
%  permeability but defects during manufacturing severely impact selectivity
%  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of
%  carbon nanotubes show promise \cite{humplik_nanostructured_2011}, but synthetic
%  techniques are unable to achieve scalable alignment and pore monodispersity
%  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm
%  pores with a narrow pore size distribution, and MD simulations of these
%  materials show that they exhibit complete rejection of solvated ions
%  \cite{murad_molecular_1998}- however, experimental rejection was low and
%  attributed to interstitial defects formed during membrane synthesis
%  \cite{li_desalination_2004}. Consequently, there is a need for a scalable
%  nanostructured porous membrane.

%  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
%  are suitable candidates for aqueous separation applications. LLCs are
%  amphiphilic molecules that share the characteristic ability of nanostructured
%  porous membrane materials to create highly ordered nanostructures with the
%  added benefits of low cost and synthetic techniques feasible for large-scale
%  production \cite{feng_scalable_2014}. LLC systems created by the monomer
%  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
%  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
%  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
%  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
%  wt\% added water results in formation of the lyotropic H\textsubscript{II}
%  phase. In both cases, Na-GA3C11 monomers assembles into mesophases made of
%  hexagonally packed, uniform size, cylinders with hydrophilic head groups
%  oriented inward towards the cylinder center. This LLC assembly is then
%  polymerized in situ to form a cross-linked polymer network to stabilize the
%  structure. The hydrophilic region can act as a pore for aqueous separations
%  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
%  specific separations by changing the monomer chemistry
%  \cite{resel_h2-phase_2000}.

  Research into the H\textsubscript{II} phase of polymerized LLC membranes has
  been revived in recent years. During early stages of exploration, hexagonal
  mesophases formed by Na-GA3C11 (Figure~\ref{fig:assembly}) could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. For this reason, research efforts
  focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment. In 2014, Feng et al.~showed that one can align 
  Col\textsubscript{h} mesophases using a magnetic field with subsequent
  cross-linking to lock the structure in place \cite{feng_scalable_2014}. In
  2016, Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}. Current efforts are
  focused on extending the method to the H\textsubscript{II} phase and 
  characterizing the performance of these newly aligned systems.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it was
  shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in diameter
  because the pores are too large \cite{zhou_supported_2005}. We do not yet
  understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for
  LLC systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group interacts with solutes of interest. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species at the membrane surface.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been created, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their model
  likely does not sufficiently model the chemical environment present in the real
  system since it only contains 16 total monomers. A separate calculation based
  on the estimated volume of the liquid crystal monomers proposes that each layer
  consists of seven monomers\cite{resel_structural_2000}. Our best chance to
  answer this question is by using a molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulation, as we
  present here. We will directly change the layer composition and note its effect
  on membrane structure.

@@

  Indications of strong $\pi$-$\pi$ stacking interactions in the direction
  perpendicular to the membrane plane support the existence of layers
  experimentally. $\pi$-$\pi$ stacking will only occur between the aromatic
  monomer head groups which leaves no description of what is happening in the
  monomer tail region. The tails may entangle isotropically while maintaining
  stacking order among headgroups.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase \textit{ab initio} studies
  of benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations over a sandwiched conformation
  \cite{sinnokrot_estimates_2002}. Substituted benzene rings exhibit an even
  stronger $\pi$-$\pi$ stacking attraction which favors the parallel displaced
  configuration in all cases except where the substitutions are extremely
  electron withdrawing\cite{waller_hybrid_2006,ringer_effect_2006}. In this
  study, we compare simulated X-ray diffraction patterns to experiment in order
  to deduce which stacking configuration is most likely.

@@

  \item What constitutes a pore, and what are the details of its pore architecture? \label{point:poredefinition}

@@

  While past work describes the Col\textsubscript{h} phase as dry
  ~\cite{feng_scalable_2014}, it has been suggested by experimentalists, in
  unpublished communications, that it is likely that neat monomer leaches small
  amounts of ambient water. Experimentally, achieving a hexagonal phase with a
  completely dry system has proven difficult. If neat monomer sits in ambient
  conditions, its color turns from transparent to slightly opaque and a hexagonal
  phase forms. Although we will not explore whether water is necessary for
  self-assembly, we hypothesize that the hydrogen bonding network formed by the
  water may play a role in structuring the pores and holding together the
  hexagonal phase. We can look at correlation functions and use simulated X-ray
  diffraction patterns to see if there is any meaningful structural difference
  between a ``dry'' and ``wet'' system.

@@

	\item \textit{R-$\pi$}: The location of the first is at $q_z$ = 1.7
	\AA$^{-1}$, corresponding to a real space separation of 3.7 \AA~. Previous
	work~\cite{feng_scalable_2014} attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

	\item \textit{R-helix}: A weak intensity line, located at exactly half
	the $q_z$ value of R-$\pi$ ($q_z$ = 0.85 \AA$^{-1}$), corresponds to a real
	space periodic spacing of 7.4 \AA~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis
	\cite{feng_scalable_2014}, meaning that if one traces the positions of the
	aromatic rings with a helical curve, then for each full turn in the helix, one
	will encounter two aromatic rings. For this reason we will refer to reflection
	as R-helix.

	\item \textit{R-alkanes}: A low intensity ring located at r = 1.4
	\AA$^{-1}$ marks the third major reflection of interest. The real space
	separation corresponds to 4.5 \AA~ which is characteristic of the average
	spacing between packed alkane chains \cite{mcintosh_organization_1980}. We will
	call this reflection R-alkanes.

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflection R-spots. The
	location of all spots is $\sim 37$ degrees from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems one can explain the spots
	by the tilt angle of the alkane chains with respect to the membrane 
	plane\cite{govind_simple_2001}.

	\item \textit{R-pores}: The final feature corresponds to the spacing
	and symmetry of the d\textsubscript{100} plane. This plane is geometrically
	related the distance between pores. The feature, which we named R-pores, is
	characterized by dots along the equatorial axis defined when $q_z$ = 0. The
	spacing between dots is indicative of the hexagonal symmetry of the packed
	pores. We observe the same information with higher resolution using SAXS
	(Fig.~\ref{fig:SAXS}).

@@

@@

  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system. To bypass the slow
  self-assembly process, we use Python scripts to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) A python script rotates and assembles monomers into layers with 
		hydrophilic centers. (c) We chose to stack twenty layers on top of each other to create
		pores. (d) The pores are duplicated and placed into a monoclinic unit cell.}\label{fig:python}
  \end{figure}
